Expert-In-The-Loop Causal Discovery: \\ Iterative Model Refinement Using Expert Knowledge

### 2025-02-10 paper/main.tex
Adds acknowledgements for use of Chat gpt
Improves plots

\subsection{Testing DAG Using CI Tests}
Our structure learning algorithm uses the idea of testing DAGs using CI tests.
In this section, we give an example of how this model testing works in
practice. We take the example of the MEC learned by the PC algorithm in
Figure~\ref{fig:intro}(a) using $ N=800 $. We first orient all the undirected
edges in the model based on our domain knowledge to obtain a DAG. In this case
we only have one undirected edge between Marital Status and Relationship. We 
orient the edge as Marital Status to Relationship. We then proceed to test
each of the implied CI of the model using a statistical test. We show the result
of some of these tests in Figure.

\todo[inline]{Add a table here}
% \begin{figure}
% 	\centering
% 	\begin{tabular}{llrr}
% 		\multicolumn{4}{c}{Z=Age, Sex} \\
% 		$X$ & $Y$ & \bm{Z} & p \\
% 		\hline
% 		Edct & Wrkc & 1.00  \\
% 		Occp & Wrkc & 1.00  \\
% 		Rltn & HrPW &  .99  \\
% 		Incm & Occp & .08   \\
% 		Incm & Wrkc &  .50  \\
% 		Incm & HrPW & .003  \\
% 		\hline
% 	\end{tabular}
% \end{figure}

Using these statistical tests we can determine the pair of variables that
should have edge between them. We can think of our proposed approach to be an
iterative application of this model testing approach.

@@

Avenues for future work include extending this approach to latent variables 
as encoded by acyclic directed mixed graphs (ADMGs). Further, if experts make 
mistakes that lead to inconsistencies during model construction (such as cycles),
this information could be leveraged in a more systematic way to allow backtracking
and recovery.
\bibliography{references}

>>>>

\subsection{Testing DAG Using CI Tests}
Our structure learning algorithm uses the idea of testing DAGs using CI tests.
In this section, we give an example of how this model testing works in
practice. We take the example of the MEC learned by the PC algorithm in
Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all the undirected
edges in the model based on our domain knowledge to obtain a DAG. In this case
we only have one undirected edge between \emph{Marital Status} and
\emph{Relationship} that we orient as $ \textrm{Marital Status} \rightarrow
\textrm{Relationship} $. We can now test the implied CIs of this DAG using
data. Figure~\ref{fig:ci_table} shows the result of this testing on some of the
implied CIs. If we use a p-value threshold of $ 0.05 $, the first $ 3 $ implied
CIs hold in the data whereas the remaining tests fail. One of the ways to fix
these failing tests is to add an edge between the variables $ X $ and $ Y $.
While this method tells up which variable pairs need to have an edge between
them, it does not tell anything about the orientation of the edge, and we would
need to orient it based on our domain knowledge.

\begin{figure}
	\centering
	\begin{tabular}{llrr}
		$X$ & $Y$ & $ \bm{Z} $ & p-value \\
		\hline
		Incm & MrtS &  Age, Occp, Rltn & $ 0.29 $     \\
		Incm & Sex  &  Occp, Rltn      & $ 0.21 $     \\
		Incm & Wrkc &  Occp 	       & $ 1.5e^{-3} $ \\
		Edct & MrtS & Incm, Occp, Rltn & $ 0.02 $       \\
		\hline
	\end{tabular}
	\caption{Results of testing some of implied CIs of the DAG in
		 Figure~\ref{fig:intro}(b) using the same CI test that was used for
		 learning it.}
	\label{fig:ci_table}
\end{figure}

@@

Avenues for future work include extending this approach to latent variables 
as encoded by acyclic directed mixed graphs (ADMGs). Further, if experts make 
mistakes that lead to inconsistencies during model construction (such as cycles),
this information could be leveraged in a more systematic way to allow backtracking
and recovery.

\begin{acknowledgements} 
	We acknowledge the use of ChatGPT (\url{https://chatgpt.com/}) for
	proofreading parts the manuscript.
\end{acknowledgements}

\bibliography{references}
